A Gentle Introduction to Compressed Sensing

### 2017-01-15 compressed_sensing/cs.tex
cs: intro is improved a little bit more

Let us consider the so-called ``Sample-Then-Compress Paradigm'' (Fig~\ref{fig:CSSampleCompressParadigm}).
We sample the signal $n$ times (vector in $\mathbb R^n$).
The transmission channel is narrow, thus, we lossy compress the sample to the dimension $d \ll n$ (vector in $\mathbb R^d$).
Afterwards, the received compressed signal is decompressed to previous dimension $n$.

This strategy has a number of flaws.
First, how big should $n$ be for us to recover the signal ``well enough'' and is it possible with the limitation of $d$?
Secondly, with $d \ll n$ the process seems to be somewhat wasteful: why do we actually need many $n$ samples if we anyway compress in to the dimension $d$?!
Thirdly, at which time points should we measure the signal?
Fourthly, every measurement in the sample could be exorbitant.
Thus, we want to reduce $n$ as much as it is possible.

@@

\section{Compressed Sensing idea}
The idea of Compressed Sensing is that the signal at Fig.~\ref{fig:CSSampleCompressParadigm} that we would like to transmit/acquire is often sparse (or ``approximately sparse'') signal in a certain ``basis''.
We will consider a very small example in the section \ref{sec:motivation} that demonstrates that this assumption is feasible.
Suppose that this signal is an $m$-dimensional vector, where $m$ is very large.
Then we want to find a linear basis in $\mathbb R^m$ for this vector to be $S$-sparse (i.e. to have only $S$ nonzero components).
The challenge of the problem is that we do not observe the whole signal.\footnote{
    Suppose that we do.
    Then we can use it as an orth in a basis and the signal is $1$-sparse.
    However, we cannot observe the whole signal because $m$ is very large.
    Take a look at Fig.\ref{fig:CSSampleCompressParadigm}.
    If the signal is an $m$-dimensional vector, then to find the basis we would need to take $n := m$ samples.
    But as we have already discussed we want to reduce $n$ as much as it is possible.
    Because of that we want to find a basis of interest without observing the whole signal.
}

Let us suppose that we now deal with a $S$-sparse signal (we still do not observe it, but know that it is $S$-sparse).
Our aim is to transmit the signal.
We can make measurements that are a certain transformation of the signal (for example, linear).
We want to make as few measurements as possible (to reduce $n$ at Fig.~\ref{fig:CSSampleCompressParadigm}).
Without the assumption of us possessing the basis in which the original signal is sparse the problem would have a trivial solution:
we would generally need to make as many measurements as is the dimension of a signal.\footnote{Think it through!}
But with the sparsity there is a chance that we need less measurements than the width of the transmission channel $d$
to reconstruct the original signal exactly or approximately.

>>>>

Let us consider the so-called ``Sample-Then-Compress Paradigm'' (Fig~\ref{fig:CSSampleCompressParadigm}).
We sample the signal $n$ times (vector in $\mathbb R^n$).
The transmission channel is narrow, thus, we lossy compress the sample to the dimension $d \ll n$ (vector in $\mathbb R^d$).
Afterwards, the received compressed sample is decompressed to previous dimension $n$ and we try to recover the signal.

This strategy has a number of flaws.
First, how big should $n$ be for us to recover the signal ``well enough'' and is it possible with the limitation of $d$?
Secondly, with $d \ll n$ the process seems somewhat wasteful: why do we actually need $n$ samples if we anyway compress to the dimension $d$?!
Thirdly, at which time points should we measure the signal?
Fourthly, every measurement in the sample could be exorbitant.
Thus, we want to reduce $n$ as much as it is possible.

@@

\section{Compressed Sensing idea}
\textbf{The first idea of Compressed Sensing} is that the signal at Fig.~\ref{fig:CSSampleCompressParadigm} that we would like to transmit/acquire is often sparse (or ``approximately sparse'') signal in a certain ``basis''.
We will consider a very small example in the section \ref{sec:motivation} that demonstrates that this assumption is feasible.
Suppose that this signal is an $m$-dimensional vector, where $m$ is very large.
Then we want to find a linear basis in $\mathbb R^m$ for this vector to be $S$-sparse (i.e. to have only $S$ nonzero components).
The challenge of the problem is that we do not observe the whole signal.\footnote{
    Suppose that we do.
    Then we can use it as an orth in a basis and the signal is $1$-sparse.
    However, we cannot observe the whole signal because $m$ is very large.
    Take a look at Fig.\ref{fig:CSSampleCompressParadigm}.
    If the signal is an $m$-dimensional vector, then to find the basis we would need to take $n := m$ samples.
    But as we have already discussed we want to reduce $n$ as much as it is possible.
    Because of that we want to find a basis of interest without observing the whole signal.
}

Let us suppose that we now deal with a $S$-sparse signal (we still do not observe it, but know that it is $S$-sparse).
Our aim is to transmit the signal.
\textbf{The second idea of CS} is that we can make measurements that are a certain transformation of the signal.
It means that we can not only measure some distinct components of the $m$-dimensional signal,
but also request, for example, a linear tranformation of the whole signal with some fixed vector.
We want to make as few measurements as possible (to reduce $n$ at Fig.~\ref{fig:CSSampleCompressParadigm}), because each measurement can be expensive.
Without the assumption of us possessing the basis in which the original signal is sparse the problem would have a trivial solution:
we would generally need to make as many measurements as the dimension of a signal.\footnote{Think it through!}
However, the sparsity gives a chance that we need less measurements than the width of the transmission channel $d$
to reconstruct the original signal exactly or approximately.
